\textbf{Sharp Bias-variance Tradeoffs of Hard Parameter Sharing in High-dimensional Linear Regression}

### 2020-10-21 arxiv_full_version/diff_cov.tex
arxiv full version

Unlike the previous section, we can no longer characterize the global minimum of $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies sharp bias-variance tradeoffs for any \emph{local minimizer} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Let $n_1, n_2$ denote task one  and two's sample size, respectively.
Suppose
\begin{align*}
	X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p}, \text{ and}
	X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p},
\end{align*}
where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

@@

 To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to this expectation as will be shown in equation \eqref{claim_largedev2}):
 %similar to Claim \ref{claim_pred_err}
\begin{align}
	 \exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}}
	=& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Equations \eqref{eq_bias_2task} and \eqref{eq_variance_2task} correspond to the bias and variance of HPS for two tasks, respectively.
Our main result in this section characterizes the asymptotic bias-variance limits in the high-dimensional setting.
Intuitively, the spectrum of $\hat{\Sigma}^{-1}$ (and hence its trace) not only depends on both tasks' sample sizes, but also depends on the ``alignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$.
However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
We introduce a key quantity $M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and as we show below, the trace of $\hat{\Sigma}^{-1}$ has an intricate dependence on the spectrum of $M$.

%Let $U\Lambda V^\top$ denote the SVD of $M$ and 
Let $\lambda_1, \lambda_2, \dots, \lambda_p$ denote $M$'s singular values in descending order.
Our main result is stated as follows.

\begin{theorem}\label{thm_main_RMT}
	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have that

@@

\begin{example}[Covariate shift]\label{ex_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example focuses on how varying covariate shifts impacts the \textit{variance} limit in equation \eqref{lem_cov_shift_eq}. For large enough $p$,
%in the left hand side of equation \eqref{lem_cov_shift_eq}:
\begin{align*}
	\bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} &\rightarrow \frac{1}{n_1 + n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}} \\
	&= \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}.
\end{align*}
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Hence the variance limit is determined by the spectrum of $M$. % To be clear, for this example we assume that the bias is $0$.
To illustrate the above result, suppose that half of $M$'s singular values are equal to $\lambda > 1$ and the other half are equal to $\lambda^{-1}$.
In particular, when $\lambda = 1$, there is no covariate shift.
As $\lambda$ increases, the severity of covariate shift increases.
We observe the following dichotomy.
\begin{enumerate}
	\item If $n_1 \ge n_2$, then the variance limit is smallest when there is no covariate shift.
	\item If $n_1 < n_2$, then the variance limit is largest when there is no covariate shift.
\end{enumerate}
\end{example}

We explain why the dichotomy happens. The variance limit in this example is equal to
$\frac{p}{2(n_1 + n_2)} f(\lambda)$, where
\[ f(\lambda) = {(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}. \]
Using the fact that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2}$, we can verify
%\begin{align*}
%	f(\lambda) - f(1) &= \left(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&+ \left(\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&- \frac{2(n_1 + n_2)}{n_1 + n_2 - p} \\
%	&= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
%\end{align*}
\begin{align*}
	f(\lambda) - f(1) &= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
\end{align*}
where $g(\lambda, a_1) \ge 0$.
%and can be derived from algebraic calculations (details omitted).
We claim that $a_1 \ge \frac{n_1 + n_2-p}{2(n_1 + n_2 )}$ if and only if $n_1 \ge n_2$, which explains the dichonomy. In fact, if $a_1>a_2$, then equation \eqref{eq_a12extra} gives that $a_1> \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$, and equation \eqref{eq_a12extra} gives that
\begin{align*}
 \frac{n_1}{n_1 + n_2} &> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) >\frac{1}{2}.
\end{align*}
%\begin{align*}
% \frac{n_1}{n_1 + n_2} &=a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} \\
%	&> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) =\frac{1}{2}.
%\end{align*}
This implies $n_1>n_2$. The other direction follows from similar arguments. % Similarly, if $a_1<a_2$, equations  \eqref{eq_a12extra000} and  \eqref{eq_a12extra} give that $a_1 < \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$ and $n_1<n_2$.
%Thus, we conclude that $f(\lambda) \ge f(1)$ if and only if $n_1 \ge n_2$.

@@

%For our purpose, we use a convenient linearization trick in linear algebra, that is, the SVD of a rectangular matrix $A$ is equivalent to the study of the eigendecomposition of the symmetric block matrix
%$$H(A):=\begin{pmatrix}0 & A \\ A^\top & 0\end{pmatrix},$$
%which is linear in $A$. This trick has been used in many random matrix literatures, such as \cite{Anisotropic, AEK_Gram, XYY_circular,DY20201}.
We say that $(W-z\id)^{-1}$ converges to a deterministic $p\times p$ matrix limit $R(z)$ if for any sequence of deterministic unit vectors $v\in \R^p$,
$$v^\top \left[(W-z\id)^{-1}-R(z)\right]v\to 0\ \ \ \text{when $p$ goes to infinity.
}$$

To study $W$'s resolvent, we observe that $W$ is equal to $\AF\AF^{\top}$ for a $p$ by $n_1 + n_2$ matrix
	\be\label{defn AF} \AF := (n_1+ n_2)^{-1/2} [\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top]. \ee
%}\HZ{what does this trick mean? use less technical words},
%This idea dates back at least to Girko, see e.g., the works \cite{girko1975random,girko1985spectral} and references therein.
Consider the following symmetric block matrix whose dimension is $p + n_1 + n_2$
%which is a linear function of $(Z^{(1)})$ and $(Z^{(2)})$:
%\begin{definition}[Linearizing block matrix]\label{def_linearHG}%definiton of the Green function
%We define the $(n+N)\times (n+N)$ block matrix
 \begin{equation}\label{linearize_block}
    H \define \left( {\begin{array}{*{20}c}
   0 & \AF  \\
   \AF^{\top} & 0
%   {Z^{(2)}V} & 0 & 0
   \end{array}} \right).
 \end{equation}
For this block matrix, we define its resolvent as
$$G(z) \define \left[H - \begin{pmatrix}z\id_{p\times p}&0\\ 0 & \id_{(n_1+n_2)\times (n_1+n_2)} \end{pmatrix}\right]^{-1},$$
%as the resolvent of $H$,
for any complex value $z\in \mathbb C$.
Using Schur complement formula for the inverse of a block matrix, it is not hard to verify that
	\begin{equation} \label{green2}
	  G(z) =  \left( {\begin{array}{*{20}c}
			(W- z\id)^{-1} & (W - z\id)^{-1} \AF  \\
      \AF^\top (W - z\id)^{-1} & z(\AF^\top \AF - z\id)^{-1}
		\end{array}} \right).%\quad \cal G_R:=(W^\top W - z)^{-1} ,
  \end{equation}

@@

\paragraph{Bias asymptotic limit.}
 For the bias limit in equation \eqref{lem_cov_derv_eq}, we show that it is governed by the derivative of $(W - z\id)^2$ with respect to $z$ at $z = 0$.
First, we can express the empirical bias term in equation \eqref{lem_cov_derv_eq} as %$W$
\begin{align}\label{calculate G'}
	(n_1 + n_2)^2 \hat{\Sigma}^{-1}\Sigma^{(2)}\hat{\Sigma}^{-1} = {\Sigma^{(2)}}^{-1/2} V W^{-2} V^{\top} {\Sigma^{(2)}}^{-1/2}.
\end{align}
Let $\cal G(z):=(W-z\id )^{-1}$ denote the resolvent of $W$.
Our key observation is that $\frac{\dd{\cal G(z)}}{\dd z} =  \cal G^2(z)$.
Hence, provided that the limit of $(W - z\id)^{-1}$ is $(a_1(z) \Lambda^2 + (a_2(z) - z) \id)^{-1}$ near $z = 0$, the limit of $\frac{\dd{\cal G(0)}}{\dd z}$ satisfies that
\begin{align}\label{cal G'0}
	\frac{\dd \cal G(0)}{\dd z} \to \frac{-\frac{\dd a_1(0)}{\dd z}\Lambda^2 - (\frac{\dd a_2(0)}{\dd z} - 1)\id}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.
\end{align}
To find the derivatives of $a_1(z)$ and $a_2(z)$, we take the derivatives on both sides of the system of equations \eqref{selfomega_a}.
%\begin{align*}
%	\frac{\dd a_1(z)}{\dd z} + \frac{\dd a_2(z)}{\dd z} = -\frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
%	\frac{\dd a_1(z)}{\dd z} + \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{\lambda_i^2 (a_1'(z) a_2 - a_2'(z) a_1)}{(\lambda_i^2 a_1 + a_2)^2} = -\frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^2}
%\end{align*}
Let $a_3 = - \frac{\dd a_1(0)}{\dd z}$ and $a_4 = - \frac{\dd a_2(0)}{\dd z}$.
%then taking implicit differentiation of equation \eqref{selfomega_a}
One can verify that $a_3$ and $a_4$ satisfy the self-consistent equations in \eqref{eq_a34extra} (details omitted).
Applying equation \eqref{cal G'0} to equation \eqref{calculate G'}, we obtain the asymptotic limit of the bias term.
%\begin{align}
%& n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ (X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)}) }^{-1} \Sigma_1^{1/2} w}^2 \nonumber\\
%&\approx  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\frac{a_3\Lambda^2 +(1+ a_4)\id_p}{(a_{1}\Lambda^2 + a_{2}\id)^2}V^\top \Sigma_2^{-1/2}\Sigma_1^{1/2}w= w^{\top} \Pi_\bias w, \label{calculatePibias}
%\end{align}
%where in the last step we used $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $V \Lambda^2 V^\top=M^\top M$. This concludes equation \eqref{lem_cov_derv_eq}.

As a remark, in order for $\frac{\dd \cal G(z)}{\dd z}$ to stay close to its limit at $z = 0$, we not only need to find the limit of $\cal G(0)$, but also the limit of $\cal G(z)$ within a small neighborhood of $0$.
This is why we consider $W$'s resolvent for a general $z$. %as opposed to the Stieljes transform of its empirical spectral distribution.

\paragraph{Schur complement and self-consistent equations.}
First, we consider the special case where $Z^{(1)}$ and $Z^{(2)}$ are both multivariate Gaussian random matrices.
By rotational invariance, we have that $Z^{(1)} U$ and $Z^{(2)} V$ are still multivariate Gaussian random matrices.
Next, we use the Schur complement formula to deal with the resolvent $G(z)$.
We show that $G(z)$'s diagonal entries satisfy a set of self-consistent equations in the limit, leading to equations in \eqref{selfomega_a}.
On the other hand, $G(z)$'s off-diagonal entries are approximately zero using standard concentration bounds.
Finally, we extend our result to general random matrices under the finite $\varphi$-th moment condition.
We prove an anisotropic local law using recent developments in random matrix theory \cite{erdos2017dynamical,Anisotropic}.
A complete proof of Theorem \ref{thm_main_RMT} will be given in Appendix \ref{sec pf RMTlemma}-Appendix \ref{sec entry}. %Appendix \ref{appendix RMT}.

>>>>

Unlike the previous section, we can no longer characterize the global minimum of $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies sharp bias-variance tradeoffs for any \emph{local minimizer} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Let $n_1, n_2$ denote task one  and two's sample size, respectively.
Suppose
\begin{align*}
	X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p} \text{ and }
	X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p},
\end{align*}
where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

@@

To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to this expectation as will be shown in equation \eqref{claim_largedev2}):
 %similar to Claim \ref{claim_pred_err}
\begin{align}
	 \exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}}
	=& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Equations \eqref{eq_bias_2task} and \eqref{eq_variance_2task} correspond to the bias and variance of HPS for two tasks, respectively.
Our main result in this section characterizes the asymptotic bias-variance limits in the high-dimensional setting.
Intuitively, the spectrum of $\hat{\Sigma}^{-1}$ (and hence its trace) not only depends on both tasks' sample sizes, but also depends on the ``alignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$.
However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
We introduce a key quantity $M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and as we show below, the trace of $\hat{\Sigma}^{-1}$ has an intricate dependence on the spectrum of $M$.
%Let $U\Lambda V^\top$ denote the SVD of $M$ and
Let $\lambda_1, \lambda_2, \dots, \lambda_p$ denote $M$'s singular values in descending order.
Our main result is stated as follows.

\begin{theorem}\label{thm_main_RMT}
	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have the following limits:

@@

\begin{example}[Covariate shift]\label{ex_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example focuses on how varying covariate shifts impacts the \textit{variance} limit in equation \eqref{lem_cov_shift_eq}. For large enough $p$,
%in the left hand side of equation \eqref{lem_cov_shift_eq}:
\begin{align*}
	\bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} &\rightarrow \frac{1}{n_1 + n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}
	= \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}.
\end{align*}
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Hence the variance limit depends on the spectrum of $M$. % To be clear, for this example we assume that the bias is $0$.
To illustrate the result, suppose that half of $M$'s singular values are equal to $\lambda > 1$ and the other half are equal to $\lambda^{-1}$.
In particular, when $\lambda = 1$, there is no covariate shift.
As $\lambda$ increases, the severity of covariate shift increases.
We observe the following dichotomy.
\begin{enumerate}
	\item If $n_1 \ge n_2$, then the variance limit is smallest when there is no covariate shift.
	\item If $n_1 < n_2$, then the variance limit is largest when there is no covariate shift.
\end{enumerate}
\end{example}
\noindent We explain why the dichotomy happens. The variance limit in this example is equal to
$\frac{p}{2(n_1 + n_2)} f(\lambda)$, where
\[ f(\lambda) = {(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}. \]
Using the fact that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2}$, we can verify
%\begin{align*}
%	f(\lambda) - f(1) &= \left(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&+ \left(\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1} \\
%	&- \frac{2(n_1 + n_2)}{n_1 + n_2 - p} \\
%	&= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
%\end{align*}
\begin{align*}
	f(\lambda) - f(1) &= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
\end{align*}
where $g(\lambda, a_1) \ge 0$.
%and can be derived from algebraic calculations (details omitted).
We claim that $a_1 \ge \frac{n_1 + n_2-p}{2(n_1 + n_2 )}$ if and only if $n_1 \ge n_2$, which explains the dichonomy. In fact, if $a_1>a_2$, then equation \eqref{eq_a12extra} gives that $a_1> \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$, and equation \eqref{eq_a12extra} gives that
\begin{align*}
 \frac{n_1}{n_1 + n_2} &> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) >\frac{1}{2}.
\end{align*}
%\begin{align*}
% \frac{n_1}{n_1 + n_2} &=a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} \\
%	&> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) =\frac{1}{2}.
%\end{align*}
This implies $n_1>n_2$. The other direction follows from similar arguments. % Similarly, if $a_1<a_2$, equations  \eqref{eq_a12extra000} and  \eqref{eq_a12extra} give that $a_1 < \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$ and $n_1<n_2$.
%Thus, we conclude that $f(\lambda) \ge f(1)$ if and only if $n_1 \ge n_2$.

@@

%For our purpose, we use a convenient linearization trick in linear algebra, that is, the SVD of a rectangular matrix $A$ is equivalent to the study of the eigendecomposition of the symmetric block matrix
%$$H(A):=\begin{pmatrix}0 & A \\ A^\top & 0\end{pmatrix},$$
%which is linear in $A$. This trick has been used in many random matrix literatures, such as \cite{Anisotropic, AEK_Gram, XYY_circular,DY20201}.
We say that $(W-z\id)^{-1}$ converges to a deterministic $p\times p$ matrix limit $R(z)$ if for any sequence of deterministic unit vectors $v\in \R^p$,
$$v^\top \left[(W-z\id)^{-1}-R(z)\right]v\to 0\ \ \ \text{when $p$ goes to infinity.
}$$
To study $W$'s resolvent, we observe that $W$ is equal to $\AF\AF^{\top}$ for a $p$ by $n_1 + n_2$ matrix
	\be\label{defn AF} \AF := (n_1+ n_2)^{-1/2} [\Lambda U^\top (Z^{(1)})^\top,V^\top (Z^{(2)})^\top]. \ee
%}\HZ{what does this trick mean? use less technical words},
%This idea dates back at least to Girko, see e.g., the works \cite{girko1975random,girko1985spectral} and references therein.
Consider the following symmetric block matrix whose dimension is $p + n_1 + n_2$
%which is a linear function of $(Z^{(1)})$ and $(Z^{(2)})$:
%\begin{definition}[Linearizing block matrix]\label{def_linearHG}%definiton of the Green function
%We define the $(n+N)\times (n+N)$ block matrix
 \begin{equation}\label{linearize_block}
    H \define \left( {\begin{array}{*{20}c}
   0 & \AF  \\
   \AF^{\top} & 0
%   {Z^{(2)}V} & 0 & 0
   \end{array}} \right).
 \end{equation}
For this block matrix, we define its resolvent as
$$G(z) \define \left[H - \begin{pmatrix}z\id_{p\times p}&0\\ 0 & \id_{(n_1+n_2)\times (n_1+n_2)} \end{pmatrix}\right]^{-1},$$
%as the resolvent of $H$,
for any complex value $z\in \mathbb C$.
Using Schur complement formula for the inverse of a block matrix, it is not hard to verify that
	\begin{equation} \label{green2}
	  G(z) =  \left( {\begin{array}{*{20}c}
			(W- z\id)^{-1} & (W - z\id)^{-1} \AF  \\
      \AF^\top (W - z\id)^{-1} & z(\AF^\top \AF - z\id)^{-1}
		\end{array}} \right).%\quad \cal G_R:=(W^\top W - z)^{-1} ,
  \end{equation}

@@

\paragraph{Bias asymptotic limit.}
 For the bias limit in equation \eqref{lem_cov_derv_eq}, we show that it is governed by the derivative of $(W - z\id)^2$ with respect to $z$ at $z = 0$.
First, we can express the empirical bias term in equation \eqref{lem_cov_derv_eq} as %$W$
\begin{align}\label{calculate G'}
	(n_1 + n_2)^2 \hat{\Sigma}^{-1}\Sigma^{(2)}\hat{\Sigma}^{-1} = {\Sigma^{(2)}}^{-1/2} V W^{-2} V^{\top} {\Sigma^{(2)}}^{-1/2}.
\end{align}
Let $\cal G(z):=(W-z\id )^{-1}$ denote the resolvent of $W$.
Our key observation is that $\frac{\dd{\cal G(z)}}{\dd z} =  \cal G^2(z)$.
Hence, provided that the limit of $(W - z\id)^{-1}$ is $(a_1(z) \Lambda^2 + (a_2(z) - z) \id)^{-1}$ near $z = 0$, the limit of $\frac{\dd{\cal G(0)}}{\dd z}$ satisfies
\begin{align}\label{cal G'0}
	\frac{\dd \cal G(0)}{\dd z} \to \frac{-\frac{\dd a_1(0)}{\dd z}\Lambda^2 - (\frac{\dd a_2(0)}{\dd z} - 1)\id}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.
\end{align}
To find the derivatives of $a_1(z)$ and $a_2(z)$, we take the derivatives on both sides of the system of equations \eqref{selfomega_a}.
%\begin{align*}
%	\frac{\dd a_1(z)}{\dd z} + \frac{\dd a_2(z)}{\dd z} = -\frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
%	\frac{\dd a_1(z)}{\dd z} + \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{\lambda_i^2 (a_1'(z) a_2 - a_2'(z) a_1)}{(\lambda_i^2 a_1 + a_2)^2} = -\frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^2}
%\end{align*}
Let $a_3 = - \frac{\dd a_1(0)}{\dd z}$ and $a_4 = - \frac{\dd a_2(0)}{\dd z}$.
%then taking implicit differentiation of equation \eqref{selfomega_a}
One can verify that $a_3$ and $a_4$ satisfy the self-consistent equations in \eqref{eq_a34extra} (details omitted).
Applying equation \eqref{cal G'0} to equation \eqref{calculate G'}, we obtain the bias limit.
%\begin{align}
%& n^2\bignorm{\Sigma_2^{1/2} \bigbrace{ (X^{(1)})^{\top}(X^{(1)}) + (X^{(2)})^{\top}(X^{(2)}) }^{-1} \Sigma_1^{1/2} w}^2 \nonumber\\
%&\approx  w^\top \Sigma_1^{1/2}\Sigma_2^{-1/2}V\frac{a_3\Lambda^2 +(1+ a_4)\id_p}{(a_{1}\Lambda^2 + a_{2}\id)^2}V^\top \Sigma_2^{-1/2}\Sigma_1^{1/2}w= w^{\top} \Pi_\bias w, \label{calculatePibias}
%\end{align}
%where in the last step we used $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $V \Lambda^2 V^\top=M^\top M$. This concludes equation \eqref{lem_cov_derv_eq}.

As a remark, in order for $\frac{\dd \cal G(z)}{\dd z}$ to stay close to its limit at $z = 0$, we not only need to find the limit of $\cal G(0)$, but also the limit of $\cal G(z)$ within a small neighborhood of $0$.
This is why we consider $W$'s resolvent for a general $z$ (as opposed to the Stieljes transform of its empirical spectral distribution discussed earlier).

\paragraph{Schur complement and self-consistent equations.}
First, we consider the special case where $Z^{(1)}$ and $Z^{(2)}$ are both multivariate Gaussian random matrices.
By rotational invariance, we have that $Z^{(1)} U$ and $Z^{(2)} V$ are still multivariate Gaussian random matrices.
Next, we use the Schur complement formula to deal with the resolvent $G(z)$.
We show that $G(z)$'s diagonal entries satisfy a set of self-consistent equations in the limit, leading to equations in \eqref{selfomega_a}.
On the other hand, $G(z)$'s off-diagonal entries are approximately zero using standard concentration bounds.
Finally, we extend our result to general random matrices under the finite $\varphi$-th moment condition.
We prove an anisotropic local law using recent developments in random matrix theory \cite{erdos2017dynamical,Anisotropic}.
The proof of Theorem \ref{thm_main_RMT} is shown in Appendix \ref{appendix RMT}. %Appendix \ref{appendix RMT}.

### 2020-10-21 arxiv_full_version/experiments_text.tex
arxiv version
arxiv full version

\section{Further Studies on Text Classification Tasks}\label{sec_text}

@@

These examples and the one in Figure \ref{fig_size} suggest a natural progressive training schedule, where we add samples progressively until performance drops.
Concretely, here is one implementation of this idea.
%In particular, Figure \ref{fig_size} (and our analysis) shows that $L(\hat{\beta}_t^{\MTL})$ behaves as a quadratic function over $\rho_1$.
%More generally, depending on how large $\Psi(\beta_1, \beta_2)$ is, $L(\hat{\beta}_t^{\MTL})$ may also be monotonically increasing or decreasing.
%Based on this observation, we propose a progressive training schedule to improve the compuational efficiency of hard parameter sharing.
\begin{itemize}
	\item We divide the training data into $S$ batches.
	We divide the training procedure into $S$ stages. During every stage, we progressively add one more data batch.
	\item During every stage, we train for $T$ epochs using only the $S$ batches. If the validation accuracy drops compared to the previous round's result or reach a desired threshold $\tau$, we terminate.
%	Algorithm \ref{alg_inc_train} in Appendix \ref{app_experiments} describes the procedure in detail.
\end{itemize}
For example, if we apply this procedure to the settings of Figure \ref{fig_ab_data} and \ref{fig_size}, it will terminate once reaching the optimal sample ratio.
%See Algorithm \ref{alg_inc_train} for a complete description for two tasks.
The advantage of this procedure is that it reduces the computational cost compared to standard round-robin training schedules.
For example, if the procedure terminates at 30\% of all batches, then SGD only passes over 30\% of its data, whereas standard round-robin training passes over 100\% of task one's data.

>>>>

\subsection{Further Studies on Text Classification Tasks}\label{sec_text}

@@

These examples and the one in Figure \ref{fig_size} suggest a natural progressive training schedule, where we add samples progressively until performance drops.
Concretely, here is one implementation of this idea.
%In particular, Figure \ref{fig_size} (and our analysis) shows that $L(\hat{\beta}_t^{\MTL})$ behaves as a quadratic function over $\rho_1$.
%More generally, depending on how large $\Psi(\beta_1, \beta_2)$ is, $L(\hat{\beta}_t^{\MTL})$ may also be monotonically increasing or decreasing.
%Based on this observation, we propose a progressive training schedule to improve the compuational efficiency of hard parameter sharing.
\begin{itemize}
	\item We divide the training data into $S$ batches.
	We divide the training procedure into $S$ stages. During every stage, we progressively add one more data batch.
	\item During every stage, we train for $T$ epochs using only the $S$ batches. If the validation accuracy drops compared to the previous round's result or reach a desired threshold $\tau$, we terminate.
%	Algorithm \ref{alg_inc_train} in Appendix \ref{app_experiments} describes the procedure in detail.
\end{itemize}
If we apply this procedure to the settings of Figure \ref{fig_ab_data} and \ref{fig_size}, it will terminate once reaching the optimal sample ratio.
%See Algorithm \ref{alg_inc_train} for a complete description for two tasks.
The advantage of this procedure is that it reduces the computational cost compared to standard round-robin training schedules.
For example, if the procedure terminates at 30\% of all batches, then SGD only passes over 30\% of its data, whereas standard round-robin training passes over 100\% of task one's data.
